---
title: "DI Example"
output: html_document
layout: template
filename: "DI-example.Rmd"
---

```{r setup, include=FALSE}

# Load required packages
library(dplyr)
library(ggplot2)
library(png)
library(ggimage)
library(ggwaffle) #devtools::install_github("liamgilbey/ggwaffle")
library(emojifont)
library(MetBrewer)
library(tidyverse)

```

## Explaining the Disproportionality Index - An Example

Suppose that we have a department of 500 employees. There are two groups of self-identifying minority employees, A and B. Group C is composed of the majority at 460 employees. Groups A and B make up 4% of employees at 20 employees each, which happens to mirror the Canadian population. At a glance, representation seems equal. 


```{r Plot-1, echo = FALSE}
# Create dataframe with group numbers repeated
Dept_data <- tibble(Group_count = rep("A", 20) %>% 
                      c(rep("B", 20)) %>% 
                      c(rep("C", 460)) )

# Move data to waffle object for plotting
Dept_data_waffle <- waffle_iron(Dept_data, 
                                aes_d(group = Group_count), 
                                rows = 20) %>% 
  mutate(label = fontawesome('fa-male'))

# 

## Plot symbols -----------------------------------------------

# Create color pallette 
colors <- c(met.brewer("Johnson", 5)[c(1,5)], "#808080")

# Plot as waffle   
ggplot(Dept_data_waffle, aes(x, y, colour = group)) + 
  geom_text(aes(label=label), 
            family = 'fontawesome-webfont', 
            size = 10) +
  coord_equal() + 
  scale_colour_manual(values = c(colors)) + 
  theme_waffle()  +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(), 
        text = element_text(size = 16)) + 
  guides(color=guide_legend(title="Group"))
  
```


However, if you take a closer look at how the 20 employees from group A and B are distributed across salary levels, a different story appears. Group A is equally distributed among salary level, but the majority of group B employees are stagnant at the lowest salary levels. 


Calculating DI further illustrates the differences in salary distribution. Above a DI of 1, a group's representation at a salary level is higher than the average of all employees. Below a DI of 1, a group's representation at a salary level is lower than the average of all employees

```{r Plot 2-DI Example, echo = FALSE, message = FALSE, warning = FALSE}
Group <- c(rep("A", 5), rep("B", 5))
Salary <- c("<60.0K", "60.0-69.9K", "70.0-84.9K", "85.0-99.9K", "100.0K+", "<60.0K", "60.0-69.9K", "70.0-84.9K", "85.0-99.9K", "100.0K+")
DI <- c(.75, 1, 1.25, 1, 1, 1.5, 1.5, 1, .75, .25)

Salary_di <- data.frame(Group, Salary, DI)

Salary_di$Salary <- factor(Salary_di$Salary, levels = c("<60.0K", "60.0-69.9K", "70.0-84.9K", "85.0-99.9K", "100.0K+"))

colors <- c(met.brewer("Johnson", 5)[c(1,5)])

ggplot(Salary_di) +
  geom_bar(stat = "identity",
           position = "dodge",
           aes(x = Group,
               y = DI,
               fill = Group, 
               alpha = Salary)) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"), 
    text = element_text(size = 15),
    legend.text = element_text(size = 12)
  ) +
  scale_fill_manual  (values = colors, 
  ) +
  labs(title = NULL,
       x = "Group",
       y = "Disproportionality Index",
       fill = NULL) + 
  guides(fill="none") +
    geom_hline(yintercept = 1, 
             linetype = "dashed")

```